### 2017-07-13 draft/unique-linear.tex
snapshot after slight cleanup

Effects attached to single record labels, intuitively $P, Q \subseteq
\{ \READ, \WRITE \}$ but it does not make sense to distinguish
$\{\WRITE, \READ\}$ from $\{\WRITE\}$, so we take the three element lattice
\begin{align*}
  O, P, Q &::= \NOTHING \mid \READ \mid \WRITE
  & \text{ordered by } \NOTHING
  &\sqsubset \READ \sqsubset \WRITE
\end{align*}

@@

  \inferrule
  { A_1 \vdash e_1 : t_1 \\
    A_2, x : t_1 \vdash e_2 : t_2 \\
    A = A_1 \OVERRIDE A_2
  }
  { A \vdash \LET x{e_1} e_2 : t}
  \end{mathpar}
  \caption{Propagation rules}
  \label{fig:field-effects}
\end{figure}
Function closures have ``infinite'' splitting $(a :^P t)^*$ into $(a :^Q t)$.
\begin{itemize}
\item $P = \NOTHING$ propagate $Q=\NOTHING$.
\item $P = \READ$ propagate $Q = \WRITE$, no action required inside closure.
\item $P = \WRITE$ propagate $Q = \WRITE$, perform copy inside of closure.
\end{itemize}
 Propagating $\READ$ is not safe in both cases because that enables a write operation on the same
 record after the closure has been constructed. Alternatively, we could have another READONLY mode,
 which prevents subsequent write operations.

>>>>

\begin{figure}[tp]
  \begin{align*}
    e &::= x \mid \LAM xte \mid \APP ee \mid \LET xee  \mid c \mid e+e \mid \IF eee\\
      &\mid \REC{ \overline{a=e}} \mid \GET ea \mid \UPD eae \\
    O,P,Q &::= \NOTHING \mid \READ \mid \WRITE & \NOTHING & \sqsubseteq\WRITE\sqsubset \READ \\
    s,t &::= \NUM \mid s \to t \mid \REC{ \overline{a :^P t}} \\
  \end{align*}
  \caption{Syntax}
  \label{fig:syntax-3}
\end{figure}
In the syntax in Figure~\ref{fig:syntax-3}, effects are attached to single record labels, intuitively $P, Q \subseteq
\{ \READ, \WRITE \}$ but it does not make sense to distinguish
$\{\WRITE, \READ\}$ from $\{\WRITE\}$, so we take the three element lattice as indicated.

@@

  \inferrule
  { A_1 \vdash e_1 : t_1 \\
    A_2, x : t_1 \vdash e_2 : t_2 \\
    A = A_1 \OVERRIDE A_2
  }
  { A \vdash \LET x{e_1} e_2 : t}
  \end{mathpar}
  \caption{Typing rules}
  \label{fig:field-effects}
\end{figure}
The novel ingredient in the typing rules are the propagation and splitting rules for the effect annotations on
record labels.
Function closures have ``infinite'' splitting $(a :^P t)^*$ into $(a :^Q t)$.
\begin{itemize}
\item $P = \NOTHING$ propagate $Q=\NOTHING$.
\item $P = \READ$ propagate $Q = \WRITE$, no action required inside closure.
\item $P = \WRITE$ propagate $Q = \WRITE$, perform copy inside of closure.
\end{itemize}
 Propagating $\READ$ is not safe in both cases because that enables a write operation on the same
 record after the closure has been constructed. Alternatively, we could have another READONLY mode,
 which prevents subsequent write operations.
